---
title: "Proving the Central Limit Theorem"
output: pdf_document
---

#Proving the Central Limit Theorem
dchen71

Overview

This report will compare the expontential distrbution against the normalized central limit theorem.

    Title (give an appropriate title) and Author Name
    Overview: In a few (2-3) sentences explain what is going to be reported on.
    Simulations: Include English explanations of the simulations you ran, with the accompanying R code. Your explanations should make clear what the R code accomplishes.
    Sample Mean versus Theoretical Mean: Include figures with titles. In the figures, highlight the means you are comparing. Include text that explains the figures and what is shown on them, and provides appropriate numbers.
    Sample Variance versus Theoretical Variance: Include figures (output from R) with titles. Highlight the variances you are comparing. Include text that explains your understanding of the differences of the variances.
    Distribution: Via figures and text, explain how one can tell the distribution is approximately normal. 

##Simulations

The following code is the code to create the simulation data. The code will setup an expontential distribution as well as its assocaited means and variability.

```{r}
library(ggplot2)

set.seed(20)

lambda = .2

sample_mean = 0
sample_var = 0
for(i in 1:1000){
    sample_mean[i] = mean(rexp(40,lambda))
    sample_var[i] = var(rexp(40,lambda))    
}

sample_mean = as.data.frame(sample_mean)

sample.mean = mean(sample_mean)
sample.var = var(sample_var)
```

Sample Means vs Theoretial Mean

```{r, echo=FALSE}
ggplot(sample_mean, aes(x=sample_mean)) + 
    geom_density() +
    geom_vline(xintercept=sample.mean, size = 1, color = 'red') 
```

The above figure shows the distribution of sampled means against the population mean. 

```{r, echo=FALSE}
hist(sample_var, col="gray", main="Sampling Distribution of the Variance", xlab="Sample Variance")
```
