---
title: "mtcars Analysis"
output: html_document
---

##Overview

The data from mtcars will be looked at and analyzed to determine the relationship between the miles per gallon and whether or not the cars are automatic or manual.

```{r}
library(ggplot2)
require(gridExtra)

auto = subset(mtcars, mtcars$am == 0)
man = subset(mtcars, mtcars$am == 1)

mtcars
```

##Exploratory Analysis

```{r, echo=FALSE}
ggplot(mtcars, aes(x=am, y=mpg)) + geom_point() + labs(title='Transmission vs MPG', x='Transmission(Auto=0, Man = 1)')

```

As can be seen in the figure above, manual cars appear to have a higher maximum mpg and minimum mpg. The range in mpg widely varies amongst  manual cars compared to automatic.

```{r}
t.test(man$mpg, auto$mpg)
```

As can be seen int two item T test of both the manual and automatic transmissions, there is a 95% confidence interval of 3.21-11.28 showing that there is likely to be some difference between manual and automatic transmissions. The p value is low enough to reject the notion that there is no difference between mpg amongst transmissions.


```{r, echo=FALSE}
p1 = ggplot(auto, aes(x=cyl, y=mpg)) + geom_point() + labs(title='Cylinders vs MPG(Automatic)')
p2 = ggplot(man, aes(x=cyl, y=mpg)) + geom_point() + labs(title='Cylinders vs MPG(Manual)')
grid.arrange(p1,p2)
```

In addition, it appears that the milage is highly dependent on the cylinders in the engine that a car posses in both types of transmissions.

##Regression modeling
```{r, echo=FALSE}

```

##Summary
