---
title: "mtcars Analysis"
output: html_document
---

##Overview

The data from mtcars will be looked at and analyzed to determine the relationship between the miles per gallon and whether or not the cars are automatic or manual.

##Exploratory Analysis

The data is first loaded and subseted/converted to factors
```{r}
library(ggplot2)
require(gridExtra)

data(mtcars)

mtcars$transmission = as.factor(mtcars$am)
mtcars$cylinders = as.factor(mtcars$cyl)
mtcars$gears = as.factor(mtcars$gear)
mtcars$carburetors = as.factor(mtcars$carb)
mtcars$vs = as.factor(mtcars$vs)

auto = subset(mtcars, mtcars$am == 0)
man = subset(mtcars, mtcars$am == 1)
```

A quick plot is setup in order to see how the data is distributed amonst transmission and mpg
```{r, echo=FALSE}
ggplot(mtcars, aes(x=am, y=mpg)) + geom_point() + labs(title='Transmission vs MPG', x='Transmission(Auto=0, Man = 1)')

```

As can be seen in the figure above, manual cars appear to have a higher maximum mpg and minimum mpg. The range in mpg widely varies amongst  manual cars compared to automatic.

```{r}
t.test(man$mpg, auto$mpg)
```

In addition, as can be seen in the two item T test of both the manual and automatic transmissions, there is a 95% confidence interval of 3.21-11.28 showing that there is likely to be some difference between manual and automatic transmissions. The p value is low enough to reject the notion that there is no difference between mpg amongst transmissions and that manual is better for mpg than automatic.

##Regression modeling
```{r}
model1 = lm(mpg ~ ., data=mtcars)
summary(model1)
```

A preliminary linear model on the mtcars dataset shows that mpg is not highly dependent on any of the independent variables. However, the coffecient of am is about 2.52023 which means that assuming all other variables remain constant, as the value goes higher(car is manual), there is a greater effect on the mpg. 


##Summary

