---
title: "Predicting the method of exercise"
output: html_document
---

#Proving the Central Limit Theorem
dchen71

Overview

This report will show off a model which will predict the manner in which a person lifted barbells using accelerometer data from indiviauls working out

##Data

First the data will be loaded.

```{r}
library(ggplot2)

train = read.csv('pml-training.csv', na.strings = c("", "NA"))
test = read.csv('pml-testing.csv', na.strings = c("", "NA"))
```

###Preprocessing

The data will first need to be cleaned up. The variables containing entries, names, as well as time and trials are removed.

```{r}
train = train[,-1:-7]
test = test[,-1:-7]
```


###Exploratory Analysis

The goal of the model is to be able to predict which of the classes that a person's method of lifting barbells correspond to. As can be seen in the table below, there are five classes with the majority of methodology falling under category A.

```{r}
table(train$classe)

```



##Model Training

```{r, echo=FALSE}


```

##Model Validation

##Prediction

