---
title: "Predicting the method of exercise"
output: html_document
---

#Proving the Central Limit Theorem
dchen71

Overview

This report will show off a model which will predict the manner in which a person lifted barbells using accelerometer data from indiviauls working out

##Data

First the data will be loaded.

```{r}
#Load data and convert blanks, divide by 0, into NA
train = read.csv('pml-training.csv', na.strings = c("", "NA", '#DIV/0!'))
test = read.csv('pml-testing.csv', na.strings = c("", "NA",'#DIV/0!'))
```

###Preprocessing

The data will first need to be cleaned up. The variables containing entries, names, as well as time and trials are removed.

```{r}
#Remove Descriptive variables
train = train[,-1:-7]
test = test[,-1:-7]

#Remove all columns but the actual gyros data with > 0.5 NAs(remove the min/max features already in data)
train = train[, colSums(is.na(train)) < nrow(train) * 0.5]
test$classe = 'dummy'
test = test[,names(train)]
```

Before a model can be trained and tested, the data from train is split up in a 60/40 ratio for model training and validation using the caret package.

```{r, echo=FALSE}
library(caret)

#Set seed for consistent partitioning
set.seed(100)

part = createDataPartition(train$classe, p = 0.6, list = FALSE)
trainset = train[part,]
testset = train[-part,]
```

###Exploratory Analysis

The goal of the model is to be able to predict which of the classes that a person's method of lifting barbells correspond to. As can be seen in the table below, there are five classes with the majority of methodology falling under category A.

```{r}
table(train$classe)

```



##Model Training



##Model Validation

##Prediction

